---
title: "team"
author: "Eric Friesth, Aashwin Lamsal"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Processing the data

```{r}
library(tidyverse)
df = read.table('diabetes.txt', header=TRUE)

df[df==""]<-NA #replacing empty strings with NA's

df = droplevels(df) #using droplevels as specified in the spec


removed_cols <- c("id","bp.2s", "bp.2d")
diabetes_reduced <- df %>% select(-removed_cols) #dropping the id, bp.2s, and bp.2d columns
#view(diabetes_reduced)

index.na=apply(is.na(diabetes_reduced), 1, any) # used to identify rows with missing value.

diabetes_clean<- na.omit(diabetes_reduced) #na.omit drops all rows with any NA Values

```
5. The way that we check if step four was done correctly, we used sum and is.na to see the
number of rows with NA values, as seen below.
```{r}
print(sum(is.na(diabetes_clean)))

```

## Exploring and Transforming data

6. You can fix the right skewness of glyhb by setting the x limits of the plot using + xlim(),
the downside of using this is we ignore outliers that lie to the right of the distribution
thus removing 44 rows.

```{r}
library(dplyr)

ggplot(diabetes_clean, aes(x = glyhb)) + geom_histogram() + labs(title = "Distribution of Glycosolated Hemoglobin", x = "Glycosolated Hemoglobin", y = "Frequency") # plotted a histogram of the variable glyhb to see the skewness

ggplot(diabetes_clean, aes(x = glyhb)) + geom_histogram() + xlim(2, 8) + labs(title = "Distribution of Glycosolated Hemoglobin", x = "Glycosolated Hemoglobin", y = "Frequency") # set limits on the x axis to create a better looking distribution without the outliers greater than 8

diabetes_clean$glyhb[diabetes_clean$glyhb > 8] <- NA # applies NA to any glyhb being greater than 8 

diabetes_clean = na.omit(diabetes_clean) # dropped those new 44 rows 

names(diabetes_clean)[names(diabetes_clean) == 'glyhb'] <- 'glyhb_star' # renamed that column to be glyhb_star

ggplot(diabetes_clean, aes(x = glyhb_star)) + geom_histogram() + labs(title = "Distribution of Glycosolated Hemoglobin", x = "Glycosolated Hemoglobin", y = "Frequency")
```

Summary Statistics of glyhb_star

```{r}

summary(diabetes_clean$glyhb_star) #summary of the variable glyhb_star

diabetes_clean %>% group_by(frame)%>% summarise(mean(glyhb_star)) # looking at the mean values of glyhb_star by the frames (small medium and large)

```

## Visualizations

```{r}
library(ggplot2)

diabetes_clean %>% group_by(frame,location) %>% summarise (mean.glyhb_star= mean(glyhb_star)) 

```

# glyhb_star against ratio
```{r}

ggplot(diabetes_clean, aes(glyhb_star, ratio, color = frame)) + geom_point() + labs(x= "Glycosolated Hemoglobin", y="Ratio", title = "Relationship between Glycosolated Hemoglobin and ratio")

```

# glyhb_star against bp.1s
```{r}
ggplot(diabetes_clean, aes(glyhb_star, bp.1s, color = frame)) + geom_point() + labs(x= "Glycosolated Hemoglobin", y="bp.1s", title = "Relationship between Glycosolated Hemoglobin and bp.1s")


```
# glyhb_star against age
```{r}
ggplot(diabetes_clean, aes(glyhb_star, age, color = frame)) + geom_point() + labs(x= "Glycosolated Hemoglobin", y="Age", title = "Relationship between Glycosolated Hemoglobin and age")


```
# glyhb_star against gender
```{r}

ggplot(diabetes_clean, aes(gender, glyhb_star, color = frame)) + geom_boxplot() + labs(x= "Gender", y="Glycosolated Hemoglobin", title = "Relationship between Glycosolated Hemoglobin and gender")

```
# glyhb_star against hip
```{r}
ggplot(diabetes_clean, aes(glyhb_star, hip, color = frame)) + geom_point() + labs(x= "Glycosolated Hemoglobin", y="Hip", title = "Relationship between Glycosolated Hemoglobin and hip")


```
# glyhb_star against weight
```{r}
ggplot(diabetes_clean, aes(glyhb_star, weight, color = frame)) + geom_point() + labs(x= "Glycosolated Hemoglobin", y="Weight", title = "Relationship between Glycosolated Hemoglobin and weight")


```

## Messy Data

13. The function gather() is used to make tables more narrow in width and longer in length
while the function spread() is used to do the opposite with a dataset, make it longer in width
and shorter in length. More specifically, these functions are used to reformat the dataset to
view it in a different way (stacking).

14. These functions are related in that they both reformat your data to make it more readable.
They complement each other on the common Key and Value. 

## Regression Models

```{r}

fit = lm(glyhb_star ~stab.glu + age + waist + ratio+ factor(frame),data=diabetes_clean)

summary(fit)

```

15. Judging from the results of our model, we an see that the mean value of the glycosolated
hemoglobin (abbreviated to GH from this point forward) is 2.62 approximately, meaning that the
level of GH is 2.62 when the stabilized glucose, age, waist, medium frame, small frame, and
ratio values are equal to 0. The 0.0079 estimate value for stabilized glucose shows that when
the value of the stabilized glucose level increases by a measure of 1, the mean GH level will
increase by 0.79%. Similarly, the estimates of 0.013, 0.018, 0.089, -0.147, and -0.276 for
age, waist, ratio, medium frame, and small frame respectively goes to show the varying changes
of the mean value of GH based on those predictors respectively, while keeping every other
predictor constant.

  With regard to the adjusted R squared, the value that tells us how much the change in our
response variable is explained by the predictors by measuring the proportion of variability in
the response that can be explained using the predictors, and the F-statistic , the value that
checks the existence of some sort of relationship (linear or logistic, in this case linear) be
between the response and predictor variables, we see that the adjusted R squared value is a
bit too low to show any correlation between the GH and predictors that we used in the model.
With regard to the F-statistic and the P-value, we see that the P-value is very small and the
F-statistic being 23.79, this leads us to believe that at least one of our six predictors is
related and significant to the GH level. 

  The results from our exploratory analysis suggest that a linear model may not be the right
approach. We had a hunch early on that this would be the case while making the scatter plots
showing the correlation of this dataset's variables. The scatter plots showed very little, if
any, correlation so a linear model wouldn't be a good approach. 

16. Interpretting the coefficients shows us the change in the mean value of GH based on an
increase of 1 unit in a predictor while all others remain constant (explained below):
  - 0.7% from an increase of 1 in stablized glucose, with all other predictors being constant
  - 1.3% from an increase of 1 in age, with all other predictors being constant
  - 1.8% from an increase of 1 in waist, with all other predictors being constant
  - 8.9% from an increase of 1 in ratio, with all other predictors being constant
  - (-14.75)% from an increase of 1 in medium frame, with all other predictors being constant
  - (-27.64)% from an increase of 1 in small frame, with all other predictors being constant


17. We can interpet the y hat value as the predicted value of y, our response variable, in a regression equation. Running the following code block shows the predicted y-values:

```{r}
view(fit$fitted.values)
```
These estimated fitted values are the estimates of the true values of glyhb_star, glycosolated hemoglobin levels. Now, using the fitted model to predict our y hat value when stab.glu = 90, age = 35, waist = 30, ratio = 5.1, and frame = small, we would end up plugging these numbers into the model to get something as such
  
  - y_hat = 2.63230 + (90).007921 + (35).013023 + (30).018357 + (5.1).089187 + (-0.276407) => 4.5301517
  
  This falls in line with glyhb_star values of records who have their frames equal to small, and shows that the fitted model might be better at predicting those who have smaller frames than those who don't.
  
18. Question 17 was a perfect example of showing the differences between predictions and inferences. Above, we predicted the glyhb_star values of a person with the following values, stab.glu = 90, age = 35, waist = 30, ratio = 5.1, and frame = small. Based on our model, we predicted that this particular person would have a glyhb_star value of 4.5301517. The lines that we wrote underneath our predicted value of glyhb_star was information that we inferred based on the prediction, which goes to show the difference between predictions and inferences. 
  
  




